% !TEX program = xelatex
%==============================================================================
% Muhammad Shaheer Luqman — Resume (single column)
% Theme matched to portfolio: accent blue #3f63ad, heading navy #1a2744,
% body slate #5a6a8a. Font: Poppins.
% IMPORTANT: Compile with xelatex (uses system Poppins font via fontspec):
%   xelatex Shaheer_Resume.tex
%==============================================================================
\documentclass[10pt,a4paper]{article}

\usepackage{amssymb}
\usepackage{fontspec}
% Poppins — loaded by file from the local fonts/ folder so the resume compiles
% on any machine without a system-wide font install. Geometric sans, slightly
% heavier than Arial.
\setmainfont{Poppins}[
  Path = fonts/,
  Extension = .ttf,
  UprightFont = Poppins-Light,
  ItalicFont  = Poppins-LightItalic,
  BoldFont    = Poppins-SemiBold,
  BoldItalicFont = Poppins-SemiBoldItalic,
  Scale = 0.90
]
\usepackage[top=1.1cm,bottom=1.1cm,left=1.4cm,right=1.4cm]{geometry}
\usepackage[dvipsnames,table]{xcolor}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fontawesome5}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{array}
\usepackage{parskip}
\usepackage{etoolbox}
\usepackage{multicol}
\setlength{\multicolsep}{2pt}
\setlength{\columnsep}{1.2em}
% Tight paragraph spacing to compress the document to two pages.
\setlength{\parskip}{1pt plus 1pt minus 1pt}

% Consistent inter-line spacing across the whole document.
% Fixed leading so every multi-line paragraph (descriptions, skills, research
% interests, etc.) wraps with the same line spacing regardless of font size.
\usepackage{setspace}
\setstretch{1.0}

%------------------------------- Colors --------------------------------------
\definecolor{accent}{HTML}{3F63AD}   % portfolio primary blue
\definecolor{accentdk}{HTML}{2E4A84} % darker blue (hover state)
\definecolor{heading}{HTML}{1A2744}  % navy headings
\definecolor{body}{HTML}{4A5878}     % slate body text
\definecolor{muted}{HTML}{7A89A8}    % muted / dates
\definecolor{rule}{HTML}{3F63AD}
\definecolor{chipbg}{HTML}{E8EEF7}   % tech-tag background

%------------------------------- Hyperlinks ----------------------------------
\hypersetup{
  colorlinks=true,
  urlcolor=accent,
  linkcolor=accent,
  pdftitle={Muhammad Shaheer Luqman — Resume},
  pdfauthor={Muhammad Shaheer Luqman}
}
% Underline + italicize hyperlink text (in addition to the accent color).
\usepackage[normalem]{ulem}
\let\orighref\href
\renewcommand{\href}[2]{\orighref{#1}{\uline{\textit{#2}}}}

%------------------------------- Section style -------------------------------
\titleformat{\section}
  {\Large\bfseries\color{heading}}
  {}{0em}
  {}[\vspace{2pt}{\color{rule}\titlerule[1.2pt]}]
\titlespacing*{\section}{0pt}{8pt}{6pt}

% Default body color
\AtBeginDocument{\color{body}}

%------------------------------- Helper macros -------------------------------
% Experience / entry header: title (left, bold navy) + date (right, muted)
% #1 = title, #2 = date, #3 = subtitle (accent italic), #4 = meta (muted small)
% Pass empty {} for any unused field.
% Title wraps within the left column; date stays flush right with a guaranteed gap.
\newlength{\datew}
\newcommand{\entryline}[2]{%
  \noindent
  \begingroup
  \settowidth{\datew}{\color{muted}#2}%
  \parbox[t]{\dimexpr\linewidth-\datew-1em\relax}{\raggedright\bfseries\color{heading}#1}%
  \hfill
  \parbox[t]{\datew}{\raggedleft\color{muted}#2}%
  \par
  \endgroup
}
\newcommand{\subline}[1]{\ifblank{#1}{}{{\color{accent}\itshape #1}\par}}
\newcommand{\metaline}[1]{\ifblank{#1}{}{{\color{muted}\small #1}\par}}
% Consistent gap before every entry across all sections. \entry ends the
% previous paragraph, removes the inter-paragraph \parskip, and inserts a
% single fixed \entryskip so every entry boundary has identical spacing
% regardless of what content (bullets, chips, description) precedes it.
\newlength{\entryskip}\setlength{\entryskip}{1pt}
\newcommand{\entry}[4]{%
  \par\vspace{-\parskip}\vspace{\entryskip}\vspace{-3pt}%
  \entryline{#1}{#2}%
  \subline{#3}%
  \metaline{#4}%
}
% Use right after \section to cancel the leading gap on the first entry.
\newcommand{\firstentry}{\vspace{3\entryskip}}

% Bullet list tuned tight
\newlist{tight}{itemize}{1}
\setlist[tight]{leftmargin=1.2em,itemsep=0.5pt,topsep=1pt,parsep=0pt,partopsep=0pt,
  label={\color{accent}\raisebox{0ex}{\normalsize\textbullet}}}

% Tech / skill chip — \strut forces a constant box height/depth so every chip
% has the same vertical size regardless of ascenders/descenders in the text.
\newcommand{\chip}[1]{%
  \setlength{\fboxsep}{2pt}%
  \colorbox{chipbg}{\color{accentdk}\small\strut\,#1\,}\hspace{2pt}}

\setlength{\parindent}{0pt}

% Remove page numbers.
\pagestyle{empty}

%==============================================================================
\begin{document}

%------------------------------- Header --------------------------------------
{\fontsize{30}{34}\selectfont\bfseries\color{accent}Muhammad Shaheer Luqman}

\vspace{2pt}
{\color{body}Applied AI \& Software Engineer who ships end-to-end computer vision, NLP, and reinforcement learning systems for industrial clients --- from research to full-stack delivery and cloud deployment. Published researcher (2$\times$ arXiv, IEEE), ICPC Gold Medalist, and summa cum laude graduate with a 3.90/4.00 CGPA.}

\vspace{2pt}
{\color{heading}%
\faEnvelope[regular]\; Email: \href{mailto:muhammadshaheerluqman@gmail.com}{muhammadshaheerluqman@gmail.com}
\quad\textbar\quad
\faMobile*\; Phone: \href{tel:+923100124127}{+92 310 0124127}
\quad\textbar\quad
\faGithub\; GitHub: \href{https://github.com/ShaheerLuqman}{/ShaheerLuqman}
}\\[3pt]
{\color{heading}%
\faLinkedin\; LinkedIn: \href{http://linkedin.com/in/shaheerluqman}{linkedin.com/in/shaheerluqman}
\quad\textbar\quad
\faGlobe\; Portfolio: \href{https://muhammad-shaheer-portfolio.vercel.app/}{muhammad-shaheer-portfolio.vercel.app}
}


%------------------------------- Experience ----------------------------------
\section{Experience}
\firstentry
\entry{Applied Research Engineer \textnormal{\small— \textbf{\textit{Retrocausal}}}}{Jun 2025 -- Present}%
{}{US-based \textbullet\ Remote}
Building \textbf{AI systems} for industrial clients end-to-end, contributing to in-house frontend/backend
software, and researching robot manipulation and Vision-Language-Action systems.
\begin{tight}
  \item Designed and shipped an end-to-end AI solution for \textbf{BMW}, owning frontend, backend, and AI components, cutting study compilation and runs from \textbf{2--3 hours of manual work to under 15 minutes}.
  \item Shipped in-house \textbf{full-stack} features across \textbf{Vue.js} frontend and \textbf{Django} backend, from API design to UI delivery.
  \item Deployed and scaled AI models on \textbf{AWS} and \textbf{RunPod} GPU infrastructure, achieving \textbf{\textasciitilde5x faster} step-name recommendation inference while \textbf{cutting GPU compute costs by up to 90\%}.
  \item Conducted \textbf{research} on unsupervised action segmentation, temporal long-video understanding, robot manipulation, and reinforcement learning, \textbf{authoring two papers currently under review} (see \textit{Research \& Publications}).
\end{tight}
\vspace{2pt}
{\footnotesize\chip{Python}\chip{PyTorch}\chip{Computer Vision}\chip{VLMs}\chip{Full-Stack}\chip{Vue.js}\chip{Django}\chip{AWS}\chip{RunPod}\chip{Reinforcement Learning}\chip{Robotics}}

\entry{Management Trainee Officer (MTO) — AI/NLP \textnormal{\small— \textbf{\textit{Confinality}}}}{Jul 2025 -- Sep 2025}%
{}{Internship \textbullet\ Evening Shift \textbullet\ Remote}
Selected for the \textbf{MTO} trainee program, working on \textbf{NLP and speech-focused AI} workflows.
\begin{tight}
  \item Engineered \textbf{NLP} pipelines for \textbf{sentiment analysis, speech detection, and transcription}.
  \item Developed \textbf{voice generation} workflows reducing manual call handling for client operations.
  \item Iteratively improved our \textbf{AI calling agents}, raising their \textbf{call transfer rate from \textless1\% to \textasciitilde6\%}.
  \item Delivered \textbf{call-performance analysis and next-step recommendations} to clients and internal teams.
\end{tight}
\vspace{2pt}
{\footnotesize\chip{NLP}\chip{Voice Generation}\chip{Sentiment Analysis}\chip{Speech Detection}\chip{Transcription}\chip{AI Calling Agents}}

\entry{Software Engineer Intern \textnormal{\small— \textbf{\textit{Retrocausal}}}}{Mar 2024 -- Jun 2024}%
{}{Internship \textbullet\ US-based \textbullet\ Remote}
Worked on \textbf{industrial-scale AI systems} across computer vision, NLP, and LLM workflows.
\begin{tight}
  \item Built a video activity recognition pipeline using \textbf{Body Pose Estimation} and \textbf{Vision-Language Models} to automate PMTS code generation for industrial workflow optimization.
  \item Developed \textbf{LLM-based prompt engineering} systems for automated analysis of industrial process data.
  \item Generated automated \textbf{Excel reports} with graph and table visualizations, building \textbf{16 modules} in total.
\end{tight}
\vspace{2pt}
{\footnotesize\chip{Python}\chip{Django}\chip{Computer Vision}\chip{LLMs}\chip{Backend}\chip{MediaPipe}\chip{Excel Automation}}

%------------------------- Research & Publications ---------------------------
\section{Research \& Publications}
\firstentry
\entry{Enhancing Human-Likeness in Reinforcement Learning Agents via Hierarchical Macro Action Quantization}{2026}%
{arXiv:2605.30928}{Under Review}
A human-like RL framework that encodes human demonstrations into macro actions via a two-level
hierarchical macro action quantization approach (HiMAQ), applied to dexterous control of a humanoid
robot hand. On the D4RL benchmarks it outperforms the non-hierarchical MAQ baseline on human-likeness
while matching or exceeding success rates, generalizing across IQL, SAC, and RLPD.
\href{https://arxiv.org/abs/2605.30928}{View on arXiv}

\entry{Unsupervised Skeleton-Based Action Segmentation via Hierarchical Spatiotemporal Vector Quantization}{2026}%
{arXiv:2604.15196}{Under Review}
A hierarchical spatiotemporal vector quantization framework for unsupervised skeleton-based temporal
action segmentation, achieving SOTA on the HuGaDB, LARa, and BABEL benchmarks. \href{https://arxiv.org/abs/2604.15196}{View on arXiv}

\entry{Ensuring Child Safety: An IoT-Based Surveillance System for Remote Monitoring and Detection of Anomalous Behavior}{2024}%
{
\vspace{2pt}ICETST 2024 \textbullet\ IEEE Xplore}{}
Proposed an intelligent IoT-based system for remote child monitoring, combining hardware sensors,
software components, and AI-driven anomaly detection to deliver real-time safety alerts.
\href{https://ieeexplore.ieee.org/document/10737980}{View on IEEE Xplore}

\vspace{4pt}
{\color{heading}\textbf{Research Interests:}}\ {\color{body}Robotics, Computer Vision, IoT + AI Systems, Vision-Language Models.}
\clearpage
%------------------------------- Education -----------------------------------
\section{Education}
\firstentry
\entry{BS Computer Science \textnormal{\small— FAST NUCES}}{Sep 2021 -- Jun 2025}%
{}{CGPA: \textbf{3.90 / 4.00 - Summa Cum Laude Honors}}

\entry{A Levels \textnormal{\small— Generation's School}}{Sep 2019 -- Jun 2021}{}{Grades: A*, 2A}
% \entry{O Levels \textnormal{\small— Habib Public School}}{}{}{Grades: 4A*, 2A, 3B}

\entry{IELTS Academic \textnormal{\small— English Proficiency}}{Jul 2026}{}{Overall Band: \textbf{7.5 / 9} \textbullet\ Reading \textbf{9.0} \textbullet\ Listening \textbf{8.5}}


%------------------------------- Projects ------------------------------------
\section{Projects}
\firstentry
\entry{PixelForge — AI-Driven Digital Marketing Platform}{}{Final Year Project (\href{https://github.com/ShaheerLuqman/PixelForge}{GitHub})}{}
\begin{tight}
  \item Built a full-stack generative AI platform using \textbf{React}, \textbf{Flask}, and \textbf{PostgreSQL} for automated advertisement image and video generation.
  \item Developed AI pipelines combining \textbf{BLIP} captioning, \textbf{Gemini}-powered slogan generation, and diffusion-based image synthesis for product-aware marketing content.
  \item Integrated \textbf{RunwayML Gen-1} workflows for AI-generated promotional videos using image-conditioned prompt engineering.
  \item Engineered an end-to-end system enabling users to generate branded advertisements from product images with minimal manual input.
\end{tight}

\entry{LushWear — Business Operation Management}{}{Client (\href{https://github.com/ShaheerLuqman/LushWear-IMS}{GitHub})}{}
Built, sold, and actively maintained for a local business to manage their day-to-day operations; currently in production use.
\begin{tight}
  \item Engineered scalable backend APIs using \textbf{FastAPI} for inventory, product catalog, stock tracking, and search operations.
  \item Built a cross-platform desktop application with \textbf{Electron}, featuring a modern responsive UI for inventory management workflows.
  \item Integrated \textbf{Supabase} for real-time database operations, authentication, and persistent cloud-backed storage.
  \item Built end-to-end \textbf{order management} with live order tracking via \textbf{courier APIs} and two-way \textbf{Shopify} integration for storefront sync.
\end{tight}


\vspace{2pt}
{\color{body}Explore more projects on my \href{https://muhammad-shaheer-portfolio.vercel.app/}{portfolio} and \href{https://github.com/ShaheerLuqman}{GitHub}.}

%------------------------------- Skills --------------------------------------
\section{Skills}

\newcommand{\skillrow}[2]{\noindent{\color{heading}\textbf{#1}}\hspace{4pt}{\color{body}#2}\par\vspace{1pt}}

\skillrow{AI / Machine Learning}{PyTorch, Scikit-learn, Hugging Face; Computer Vision (YOLO, MediaPipe); Diffusion Models (Stable Diffusion); NLP (BERT, Whisper, ElevenLabs); Reinforcement Learning (IQL, SAC, RLPD); Robotics \& Robot Manipulation; LLM Prompt Engineering.}
\skillrow{Software Engineering}{Backend (Django, Flask, FastAPI, Express.js); Frontend (React, Next.js, Vue); REST APIs, GraphQL; Postman; Databases (MySQL, MongoDB, Oracle SQL, PostgreSQL); Cloud (AWS, RunPod); System Design.}
\skillrow{Systems \& Tools}{Selenium (Web Automation), Git \& Version Control, Linux, Docker.}
\skillrow{Programming Languages}{Python, C / C++, JavaScript.}
\skillrow{Soft Skills}{Client-Facing \& Stakeholder Communication, Attention to Detail, Ownership \& Accountability}

%--------------------------- Honors & Awards ---------------------------------
\section{Honors \& Awards}
\begin{tight}
  \item \textbf{ICPC Gold Medalist} — ICPC National Competitive Programming \textit{(Nov 2024)}
  \item \textbf{Summa Cum Laude} — Graduated with a high GPA (3.90 / 4.00), FAST NUCES \textit{(2025)}
  \item \textbf{ICPC Bronze Medalist} — ICPC National Competitive Programming \textit{(Feb 2024)}
  \item \textbf{Winner} — Code Sprint, PROCOM 2023 \textit{(2023)}
  \item \textbf{Best Team Award} — Coder's Cup, ACM FAST
  \item \textbf{4x Rector's List (4.0 GPA)} — FAST NUCES, 4 semesters
  \item \textbf{4x Dean's List (3.5+ GPA)} — FAST NUCES, 4 semesters
\end{tight}

%----------------------- Leadership & Volunteer ------------------------------
\section{Leadership \& Volunteer Experience}
\begin{tight}
  \item \textbf{Director Technical Affairs} — Developers Day'25
  \item \textbf{Technical Lead} — ACM NUCES
  \item \textbf{CS Competition Head} — Developers Day'24
  \item \textbf{CS Competition Co-Head} — PROCOM'24
  \item \textbf{Training \& Mentorship Head} — ACM NUCES
\end{tight}

\end{document}
